---
title: "Project Proposal"
author: "The Four-mula: Hellen Han, Lila Rogers, Amy Duan, Eva Aggarwal"
format: pdf
execute: 
  warning: false
  message: false
editor: visual
---

```{r}
#| label: load-pkg-data

library(tidyverse)
library(tidymodels)

imdb_data <- read_csv("data/Imdb Movie New Data_v2.csv")
```

## Introduction - Lila

-   An introduction to the subject matter you’re investigating (citing any relevant literature)

-   Statement of a well-developed research question.

-   The motivation for your research question and why it is important

-   Your team’s hypotheses regarding the research question

    -   This is a narrative about what you think regarding the research question, not formal statistical hypotheses.

## Data description - Eva

We obtained this data set from [Kaggle](https://www.kaggle.com/), an online data science platform with a collection of community-developed open data sets. Here is the [link](https://www.kaggle.com/datasets/anandshaw2001/imdb-data) to where we found the data.

This data was collected by **Anand Shaw** from the **IMDb website** using sites, and converted into a .csv file. The data was updated on a daily basis until 2 months ago.

This data set collects information available on the IMDb website for different movies, such that each observation is some characteristics of a specific movie. It measures the following **21 characteristics** per movie:

1\. `id`: A unique identifier for each movie.

2\. `title`: The name of the movie.

3\. `vote_average`: The average rating the movie has received from users (on a scale, typically from 0 to 10).

4\. `vote_count`: The total number of votes or ratings submitted for the movie.

5\. `status`: The current state of the movie (e.g., "Released," "Post-Production").

6\. `release_date`: The date when the movie was officially released.

7\. `revenue`: The total earnings the movie made (usually in USD).

8\. `runtime`: The duration of the movie in minutes.

9\. `adult`: Indicates whether the movie is classified as adult content (e.g., "True" or "False").

10\. `budget`: The total cost of producing the movie (usually in USD).

11\. `imdb_id`: The unique identifier for the movie on IMDb (Internet Movie Database).

12\. `original_language`: The language in which the movie was originally produced (e.g., "en" for English).

13\. `original_title`: The original title of the movie in its native language.

14\. `overview`: A brief summary or description of the movie's plot.

15\. `popularity`: A metric indicating how popular the movie is (typically based on views, searches, or ratings).

16\. `tagline`: A short phrase or slogan associated with the movie.

17\. `genres`: The categories or genres the movie belongs to (e.g., Action, Comedy, Drama).

18\. `production_companies`: The names of the companies involved in producing the movie.

19\. `production_countries`: The countries where the movie was produced.

20\. `spoken_languages`: The languages spoken in the movie.

21\. `keywords`: Important terms or phrases associated with the movie, often used for categorization or search.

```{r}
#| label: glimpse-data

glimpse(imdb_data)
```

Due to the large size of the original data set, for the purpose of uploading this to our repository, we dropped the last characteristic, `keyword`, from the data set due to its redundancy, irrelevance to the research question and presence of a significant number of null values. So, our data set now contains 20 variables.

## Data processing

-   Description of data processing you need to do to prepare for analysis, such as joining multiple data sets, handling missing data, etc.

For the following Imdb movies dataset, to prepare the dataset for analysis, first, we need to consider missing data values within the dataset, eliminating observations that are incomplete across our predictor/response variables and filtering out movies that haven't yet been released. The majority of these missing values appear in variables that aren't relevant to our regression analysis, however, we still need to account for missing values in our predictor variables.

Furthermore, we need to check for data inconsistencies in the quantitative variables by analyzing the summary statistics for the predictor variables and noting whether there are extreme outliers that are improbable (quantitative variables with an unreasonable value of 0 or negative values). For ease of interpretation, we might scale some quantitative variables, such as revenue and budget scaled in millions of dollars.

In addition, our categorical predictors, such as adult and original language, need to be converted to factors to allow us to use them in our regression analysis.

A possible consideration for our dataset in the data processing phase is to filter the dataset for more recent Imdb movies to gain a better and more representative understanding of recent Imdb movie revenues.

-   Visualizations, summary statistics, and narrative to describe the distribution of the [**revenue**]{.underline} variable.

```{r}
summary(imdb_data$revenue)

imdb_data_revenue <- imdb_data %>% 
  filter(revenue > 0) 
summary(imdb_data_revenue$revenue)
```

```{r}
imdb_data %>% 
  filter(revenue > 0) %>% 
  ggplot(mapping = aes(x = revenue)) + 
  geom_histogram() + 
  labs(title = "Distribution of Imdb Movie Revenue",
       x = "Total Revenue (USD)", 
       y = "Count") + 
  scale_x_continuous(limits = c(0, 100000000)) + 
  scale_y_continuous(limits = c(0, 2500)) 

imdb_data %>% 
  filter(revenue > 0) %>% 
  ggplot(mapping = aes(x = revenue)) + 
  geom_boxplot() + 
  labs(title = "Distribution of Imdb Movie Revenue",
       x = "Total Revenue (USD)", 
       y = "Count") + 
  scale_x_continuous(limits = c(0, 10000000))
```

From the summary statistics, it is observable that the majority of these revenue values are unknown values (represented by 0), as the 1st quartile, median, and 3rd quartile are all 0 dollars. Thus, to obtain representative visualizations of the distribution of the revenue response variable, we analyze Imdb movies that have a revenue greater than 0 (non-missing values in this dataset).

The visualizations describe that the distribution of Imdb movie revenue is heavily right-skewed and unimodal, indiciating that there are various outliers in Imdb movies that generate significantly higher revenues. For the purpose of creating meaningful visualizations, we omitted Imdb movie revenues that are greater than 100,000,000 dollars for the histogram and greater than 10,000,000 dollars for the boxplot. Furthermore, the distribution of Imdb movie revenues has a center of approximately 1,818,000 dollars, described by the median, and a spread of approximately 18,616,000 dollars, described by the interquartile range. There are multiple outliers in the dataset for Imdb movie revenue greater than approximately 5,000,000 dollars, as shown in the boxplot visualization, as a few Imdb movies have significantly higher generated revenues. The majority of Imdb movies generate between approximately 0 and 2,000,000 dollars in revenue.

## Analysis Approach - Hellen

Given the information provided by the dataset, we are interested in investigating into what factor contributed to the movie revenue (`revenue`) of all IMDb movies within the dataset to provide more insights into recent trends in the movie industry and how producers and directors can maximize their revenue with every movie they worked on. Some potential predictor variables of interest includes:

-   `vote_average`: The average rating the movie has received from users, on a scale from 0-10 can help us understand how popularity within users influence movie revenue

-   `status`: The current state of the movie ("Released","Post-Production" etc.) can help us understand whether the state of the movie influence revenue

-   `release_date`: The date when the movie was officially released can help us understand the trend of movie revenue throughout the time frame

-   `runtime`: The duration of the movie in minutes could also influence movie revenue

-   `adult`: whether the movie is classified as adult content can also influence movie revenue, as specific content might attract more revenue than the others

-   `budget`: The total cost of producing the movie might indicates how much upfront capital is poured to the production and thereby influence the movie revenue.

-   `original_language`: the language in which the movie was originally produced might also influence movie revenue, as specific language might cover more audience than the others and thereby leading to more revenue generation

-   `genres`: movie genre might also influence movie revenue

Given that our response variable is quantiative and that we have multiple predictor variable(both quantitative and categorial), we believe multiple linear regression model is the most suitable regression model. Because it can handle multiple predictors simultaneously and quantify the impact of each predictor on the revenue.

## Data Dictionary - Hellen
